% $  Id: collabide.tex  $
% !TEX root = main.tex

%%
\section{CollabIDE}
\label{sec:collab-ide}

CollabIDE is an integrated development environment on the cloud that lets development teams 
collaborate on a software project. CollabIDE has three key features which help reduce the overhead 
problems described previously. These features are concurrent development, contribution management 
and version management.

\subsection{Concurrent Development}
Concurrent development means that all the actions a developer does in the IDE are synchronic. This lets all developers view in real time what changes are being made by other team members in the code editor or the control panel. In figure [] we show how the IDE highlights the code belonging to a certain team member. This feature eliminates the need of constantly having to obtain the latest changes made by other team members. Additionally, conflicts become nonexistent as there will always be only one version of the content being seen by the developers at a given time. The awareness of what each team member has done increases too with this feature, boosting coordination and allowing developers to know in little time what progress has been made in a project since the last time they worked on it.   

\subsection{Contribution Management}
In a project, each developer is assigned a unique color which is used to highlight the contributions of code said developer has done. Additionally, each developer has a contribution state that determines if his contributions are visible or not to the rest of the team. In figure [] we show how this feature is integrated into the IDE. With contribution management, any developer can easily identify the contributions of each team member and toggle their visibility as needed. From a product variant point of view, this feature eases the process of building new variants from existing fragments of code, effectively reducing the overhead derived from executing initial configuration processes.

\subsection{Version Management}
There can be many product versions at a time in a project. Developers have the option of easily creating and updating product versions which are snapshots of the current contents of the code editor. In a project there is always one main version, and all the other versions are derived from his one, this could be seen as the branching system that is used in git. Thanks to the concurrent development feature, all the team members obtain a new version the moment it is created. The developers can also switch between versions and view immediately the contents of the editor at the time the version was created or edited. In figure [] we show how product versioning is managed in the IDE. Proccess related to version management must be carried out multiple times in the lifespan of a project. While these processes individually don’t need much time to execute, in the long-term, they can have a lasting impact on productivity. As the time required for managing different product versions is reduced with this feature, this long-term impact can be reduced effectively.

\subsection{Implementation}
To implement the features described previously, various contexts must be considered. Each developer 
and each product version represents a different context. To achieve this context management, we used 
the \ac{COP} paradigm~\cite{salvaneschi+12survey}, specifically the Context Traits implementation. 
Context Traits provides the necessary elements to define a specific behavior during runtime for each 
context. CollabIDE features were build making use of these elements, which helped define the 
contribution state of a developer among other elements.
For the code editor of CollabIDE we used Firepad\footnote{\url{https://firepad.io/\#1}} which is an open 
source collaborative text editor. Firepad stores all the changes made in the editor in Firebase’s real 
time non relational database. Additional to the data Firepad stores, we store user and product versions 
data in Firebase.

\endinput